---
title: "Homework 4"
author: "Sara Gonzalez"
date: "May 22, 2023"
output: word_document
execute:
  warning: false
  message: false
---
# Set-up

```{r libraries}
library(tidyverse)
library(here)
library(lterdatasampler)
library(naniar)
library(flextable)
library(car)
library(performance)
library(broom)
library(ggeffects)
```


```{r read-in-data, warning=FALSE, message=FALSE}
# organize using here package
fishes <- read_csv(here("data", "ntl6_v12.csv"))
```


```{r filtering-data}
# filter data based on what columns we need
perch_data <- fishes %>%
  filter(spname=="YELLOWPERCH") %>% 
  subset(select = c(year4, spname, length, weight))
```

# Problem 1

## Hypotheses

1.  Mathematical Hypotheses

    1. Null: The predictor variable does not predict the response variable.
    2.  Alternative: The predictor variable does predict the response variable.

    Biological Hypotheses

    1.  Null: Fish length does not significantly predict fish weight for trout perch.
    2.  Alternative: Fish length significantly predicts fish weight for trout perch.

## Missing Data

2.  

```{r missing-data-vis}
gg_miss_var(perch_data)+
  labs(title = "Missing Data", caption = "The missing data may be relevant to the hypotheses because 
       there are no missing values for length, but a lot for weight.") # added title and caption to plot
```

## Test

3.  

```{r linear-model-perch}
perch_model <- lm(weight ~ length, data = perch_data)
perch_model  # y intercept and slope for fish length
```

## Check Assumptions

4.  

```{r checking-assumptions}
# to display all four at once:
par(mfrow = c(2,2)) # set up 2 x 2 grid
plot(perch_model)
```

5.  The residual vs. fitted plot showcases the variance and tests for homoscedasticity. Since the residuals are relatively even and randomly distributed, I would say this meets the homoscedastic assumption.
    The normal qq plot tests for normal distribution. Since it is for the most part on the dotted line, I think the data meets the normal distribution assumption.
    The scale-location plot essentially says the same thing as Residuals vs Fitted, except using the square root of the standardized residuals. Again, the data looks randomly distributed, so I would say it is another way to show that the homoscedastic assumption is met.
    The residuals vs leverage plot tells you if there are any influential outliers. There are no outliers outside the dotted line on the left side, I will say none of the outliers are influential towards the model estimates.

## Results

6.  summarize

```{r summary}
summary(perch_model)
```

7.  ANOVA table

```{r table}
model_squares <- anova(perch_model) # use anova to make table
model_squares
model_squares_table <- tidy(model_squares) %>% 
  mutate(p.value = case_when(
    p.value < 0.001 ~ "< 0.001")) %>% 
  flextable() %>% 
  set_header_labels(df = "Degrees of Freedom", sumsq = "Sum of Squares", meansq = "Mean of Squares") # adding more informative titles
model_squares_table
```

8.  The ANOVA table relates to the information from the summary object in that it extract information from the summary table (p-value and degrees of freedom) and tells you relevant information about degrees of freedom, test statistic, p-value, R\^2 as well as where the latter three came from.

9.  Results Summary
    We found that our model including length significantly predicted the weight of the perch (F(1, 12771) = 49,330, p < 0.001, α = 0.05). With a 95% confidence interval, we found that for every unit increase in weight, there is a 0.83 ± 0.004 cm increase in length. 

10. Visualization

```{r pulling-out-predictions}
predictions <- ggpredict(perch_model, terms = "length")
View(predictions) # terms corresponds to whatever predictor was in model
```

```{r plotting-predictions, warning=FALSE, message=FALSE}
plot_predictions <- ggplot(data = perch_data, aes(x = length, y = weight))+
  geom_point() + # first plot underlying data
  # plotting model predictions from predictions object from ggeffects
  geom_line(data = predictions, aes(x = x, y = predicted), color = "blue", linewidth = 1) +
  geom_ribbon(data = predictions, aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  labs(x = "Length (cm)", y = "Weight (g)", title = "The Weight for Trout Perch at Different Lengths", caption = "The weights of trout perch at a given length from North Temperate Lakes LTER. Black points 
       represent fish weights at a particular length. Blue line represents the trend in which the data follows. ") +
  theme_classic()+
  theme(text = element_text(family = "Times New Roman"),
axis.title = element_text(size = 14),
  )
plot_predictions
```

link to repo: https://github.com/gnlsara/ENVS-193DS_homework-04_Gonzalez-Sara